Unity in diversity: Children from 17 diverse communities process gaze in similar ways

### 2024-07-18 paper/paper.Rmd
correct spelling mistakes, update to APA 7

---
title             : "A universal of human social cognition: Children from 17 communities process gaze in similar ways"
shorttitle        : "Gaze-following across 17 communities"

author: 
  - name          : "Manuel Bohn (ORCID: 0000-0001-6006-1348)"
    affiliation   : "1,2,*"
    corresponding : yes    # Define only one corresponding author
    address       : "Universitätsallee 1, 21335 Lüneburg, Germany"
    email         : "manuel.bohn@leuphana.de"
    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
      - "Conceptualization"
      - "Methodology"
      - "Formal Analysis"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Julia Prein (ORCID: 0000-0002-3154-6167)"
    affiliation   : "1,2,*"
    role:
      - "Conceptualization"
      - "Methodology"
      - "Software"
      - "Investigation"
      - "Writing - Review & Editing"
  - name          : "Agnes Ayikoru"
    affiliation   : "3"
    role:
      - "Investigation"
  - name          : "Florian M. Bednarski (ORCID: 0000-0003-4384-4791)"
    affiliation   : "4"
    role:
      - "Investigation"
      - "Writing - Review & Editing" 
  - name          : "Ardain Dzabatou"
    affiliation   : "5"
    role:
      - "Investigation"
  - name          : "Michael C. Frank (ORCID: 0000-0002-7551-4378)"
    affiliation   : "6"
    role:
      - "Investigation"
      - "Writing - Review & Editing"
  - name          : "Annette M. E. Henderson (ORCID: 0000-0003-4384-4791)"
    affiliation   : "4"
    role:
      - "Investigation"
      - "Writing - Review & Editing"      
  - name          : "Joan Isabella"
    affiliation   : "3"
    role:
     - "Investigation" 
  - name          : "Josefine Kalbitz"
    affiliation   : "2"
    role:
      - "Investigation"
      - "Writing - Review & Editing"
  - name          : "Patricia Kanngiesser (ORCID:0000-0003-1068-3725)"
    affiliation   : "7"
    role:
      - "Investigation"
      - "Writing - Review & Editing" 
  - name          : "Dilara Keşşafoğlu (ORCID: 0000-0002-7356-0733)"
    affiliation   : "8"
    role:
      - "Investigation"
      - "Writing - Review & Editing" 
  - name          : "Bahar Köymen (ORCID: 0000-0001-5126-8240)"
    affiliation   : "9"
    role:
      - "Investigation"
      - "Writing - Review & Editing"  
  - name          : "Maira V. Manrique-Hernandez"
    affiliation   : "2"
    role:
      - "Investigation"
  - name          : "Shirley Magazi (ORCID: 0009-0006-0479-9800)"
    affiliation   : "10"
    role:
      - "Investigation"
  - name          : "Lizbeth Mújica-Manrique"
    affiliation   : "2"
    role:
      - "Investigation"
      - "Writing - Review & Editing" 
  - name          : "Julia Ohlendorf"
    affiliation   : "2"
    role:
      - "Investigation"
  - name          : "Damilola Olaoba"
    affiliation   : "2"
    role:
     - "Investigation"
  - name          : "Wesley R. Pieters (ORCID:0000-0002-6152-249X)"
    affiliation   : "10"
    role:
      - "Investigation"
      - "Writing - Review & Editing" 
  - name          : "Sarah Pope-Caldwell"
    affiliation   : "2"
    role:
     - "Investigation"    
  - name          : "Katie Slocombe (ORCID: 0000-0002-7310-1887)"
    affiliation   : "11"
    role:
     - "Investigation"
     - "Writing - Review & Editing"
  - name          : "Robert Z. Sparks (ORCID: 0000-0001-7545-0522)"
    affiliation   : "6"
    role:
      - "Investigation"
  - name          : "Jahnavi Sunderarajan"
    affiliation   : "2"
    role:
      - "Investigation"
  - name          : "Wilson Vieira"
    affiliation   : "2"
    role:
     - "Investigation"
  - name          : "Zhen Zhang (ORCID: 0000-0001-9300-0920)"
    affiliation   : "12"
    role:
      - "Investigation"
      - "Writing - Review & Editing"   
  - name          : "Yufei Zong (ORCID: 0009-0000-5012-0244)"
    affiliation   : "12"
    role:
      - "Investigation"
  - name          : "Roman Stengelin (ORCID: 0000-0003-2212-4613)"
    affiliation   : "2,10,+"
    role:
      - "Conceptualization"
      - "Methodology"
      - "Investigation"
      - "Writing - Review & Editing" 
  - name          : "Daniel B. M. Haun (ORCID: 0000-0002-3262-645X)"
    affiliation   : "2,+"
    role:
      - "Conceptualization"
      - "Funding acquisition"
      - "Writing - Review & Editing"

@@

authornote: |
   The authors would like to thank Luke Maurits for statistical advice. Manuel Bohn was supported by a Jacobs Foundation Research Fellowship (2022-1484-00). We are gratful to thank all children and caregivers for participating in the study. We thank the Max Planck Society for the Advancement fo Science.

abstract: |
  Theoretical accounts assume that key features of human social cognition are universal. Here we focus on gaze following, the bedrock of social interactions and coordinated activities, to test this claim. In this comprehensive cross-cultural study spanning five continents and 17 distinct cultural communities, we examined the development of gaze following in early childhood. We identified key processing signatures through a computational model that assumes that participants follow an individual’s gaze by estimating a vector emanating from the eye-center through the pupil. Using a single reliable touchscreen-based task, we found these signatures in all communities, suggesting that children worldwide processed gaze in highly similar ways. Absolute differences in performance between groups are accounted for by a cross-culturally consistent relationship between children’s exposure to touchscreens and their performance in the task. These results provide strong evidence for a universal process underlying a foundational socio-cognitive ability in humans that can be reliably inferred even in the presence of cultural variation in overt behavior.

keywords          : "keywords"
wordcount         : "X"

bibliography      : "library.bib"
csl: science.csl

@@

classoption       : "man"
output            : papaja::apa6_doc
---

@@

Human socio-cognitive skills enable unique forms of communication and cooperation that provide a bedrock for cumulative culture and the formation of complex societies  [@tomasello2020adaptive; @laland2021understanding; @wellman2014making; @henrich2016secret; @tomasello2003makes; @legare2019development; @heyes2018cognitive]. The eyes are the proverbial "window to the mind" and eye gaze is essential for many social reasoning processes  [@shepherd2010following; @doherty2006development]. Others’ eye gaze is used to infer their focus of visual attention, which is a critical aspect of coordinated activities, including communication and cooperation [@tomasello2007reliance; @scaife1975capacity]. The ability to follow gaze emerges early in childhood [@tang2023slow; @gredeback2010development; @byers2021development; @del2019developmental], and individual differences in children’s gaze-following ability predict later life outcomes, most notably communicative abilities [@brooks2005development; @carpenter1998social]. Difficulties with gaze following have been linked to developmental disorders, including Autism [@itier2009neural].

There is a widely held assumption that, despite substantial variation in developmental contexts, gaze-following works and develops in the same way across human societies [@tomasello2019becoming]. However, this view is seemingly at odds with the substantial diversity in socio-cognitive development that cross-cultural studies have revealed [@miller2018contributions; @wellman2014making; @mayer2013synchrony; @dixson2018scaling; @taumoepeau2019cross]. One potential source for this paradox lies in the reliance on aggregated measures in cross-cultural studies. Absolute differences in mean performance across communities are interpreted as a signal of different underlying cognitive processes. In this study, we resolve this paradox by instead focusing on processing signatures that can be investigated independently of absolute community-level differences. This allows us to directly evaluate the empirical foundation of claims about universal features of human social cognition. To this end, we conducted a pre-registered, large-scale, cross-cultural study on the development of gaze-following abilities to study potentially universal processing signatures. These signatures were derived from a computational model that assumes that participants follow gaze by estimating a vector emanating from the eye center through the pupil [@prein2023variation].

The `r n_distinct(data$subjid)` participants who took part in the study lived in 17 different communities across 14 countries and five continents (Fig. \@ref(fig:fig1)A, Tab. \@ref(tab:tab1)). These countries represent ~46% of the world’s population. Communities covered a broad spectrum of geographical locations, social and political systems, languages, and subsistence styles (see Supplemental Materials). This diversity allowed us to overcome the common pitfall of cross-cultural studies that compare urban communities from the global north to rural communities from the global south [@barrett2020towards].

```{r fig1, include = T, fig.align = "center", fig.cap = "(A) Data collection sites. Points show the approximate geographical location of the data collection sites, coloring shows the sample sizes. (B) Screenshots from the task. Screenshots from the task. The upper scene depicts the start and the lower the choice phase in a test trial. Participants had to use the gaze of the agent to locate the balloon and touch the location on the hedge where they thought the balloon was. Agents, audio recordings and backgrounds were adapted to each cultural setting. (C) Drawings used as agents across cultural settings. (D) Graphical illustration of the cognitive model. Individuals infer the target of an agent’s attention by estimating a vector based on the position of the pupils within the eyes. This process is noisy, illustrated by the different vectors (transparent lines). Individuals differ in their level of precision (indicated by sigma). For a given level of precision, the further the target lands from the centre of the screen, the less precise the model predicts individuals to be. Solid and transparent dots show simulated means and individual data points to illustrate the predicted effect of target position.", out.width="75%"}
knitr::include_graphics("../figures/fig1.png")
```

```{r tab1, include = T}
tba1 <- data %>%
  separate(community, into = c("community", "country"), sep = " \\(")%>%
  mutate(country = str_remove(country, "\\("),
         country = str_remove(country, "\\)"))%>%
  distinct(subjid, .keep_all = T)%>%
  group_by(community, country, continent)%>%
  summarise(N = n_distinct(subjid), 
            m = sum(sex == "m", na.rm = T), 
            mean_age = mean(ageinyears),
            min_age = min(ageinyears), 
            max_age = max(ageinyears), 
            touchscreen = mean(touchscreen, na.rm = T))%>%
  mutate_if(is.numeric, format, digits = 2, nsmall = 2)%>%
  mutate(age = paste0(mean_age, " (", min_age, " - ", max_age, ")"))%>%
  mutate(N = paste0(N, " (", m,")"))%>%
  select(-c(mean_age, min_age, max_age))%>%
  arrange(continent, country)%>%
  group_by(continent) %>%
  mutate(continent = as.character(continent),
         continent = replace(continent, duplicated(continent), " ")) %>%
  ungroup() %>%
  mutate(country = replace(country, duplicated(country), " "))%>%
  mutate(language = c("Spanish (Rioplatense)", "Spanish (Mexican)", "English (American)", "Hai||om", "Khwedam", "English (Nigerian)2", "English (Nigerian)", "BaYaka", "Lingala", "Kiswahili", "Bemba", "German", "English (British)", "Mandarin", "English (Indian) / Marathi", "Turkish", "English (New Zealand)"))%>%
  select(continent, country, community, N, age, language, touchscreen)

apa_table(tba1,
          col.names = linebreak(c("Continent", "Country", "Community", "N(male)", "Age (range)", "Language", "Touchscreen exposure1")),
          caption = "Participant demographics.",
          #align = paste0("m{", 1/(ncol(tba1) + 1), "\\linewidth}"),
          longtable = TRUE,
          note = "1 Proportion of participants who have access to touchscreens according to parental questionnaire. 2 Local collaborators and piloting suggested that Nigerian English is suitable for Windhoek as well.",
          escape=TRUE)
```

We used an animated picture book tablet task in which participants had to locate a hidden object based on observing an agent’s gaze. Children watched a balloon disappear behind a hedge. An agent followed the trajectory of the balloon with their eyes (Fig. \@ref(fig:fig1)B). The key dependent variable was the (im)precision with which children located the agent’s focus of attention, that is, the deviation between where the agent looked (where the object was) and the child’s response. We adapted visuals and audio instructions specifically for each of the 17 communities. Previous work demonstrated excellent individual-level measurement properties for this task in a German sample [@prein2023tango].

@@

As the first step, we investigated developmental improvements, that is, how children become more precise at estimating the target location with age. Across all 17 communities, we found a substantial increase in average levels of precision with age (fixed effect in Bayesian regression model [@burkner2019bayesian]: $\beta$ = `r post_fix%>%filter(term == "ageinyears")%>%pull(Estimate)`, 95% HDI (`r post_fix%>%filter(term == "ageinyears")%>%pull(Q2.5)` - `r post_fix%>%filter(term == "ageinyears")%>%pull(Q97.5)`); range of community-level (random) effects: $\beta_{min}$ = `r post_agg%>%filter(type == "ageinyears")%>%arrange(mean)%>%tail(1)%>%pull(mean)`, 95% HDI (`r post_agg%>%filter(type == "ageinyears")%>%arrange(mean)%>%tail(1)%>%pull(lci)` - `r post_agg%>%filter(type == "ageinyears")%>%arrange(mean)%>%tail(1)%>%pull(uci)`) to $\beta_{max}$ = `r post_agg%>%filter(type == "ageinyears")%>%arrange(mean)%>%head(1)%>%pull(mean)`, 95% HDI (`r post_agg%>%filter(type == "ageinyears")%>%arrange(mean)%>%head(1)%>%pull(lci)` - `r post_agg%>%filter(type == "ageinyears")%>%arrange(mean)%>%head(1)%>%pull(uci)`).

>>>>

---
title             : "A universal of human social cognition: Children from 17 communities process gaze in similar ways"
shorttitle        : "gaze following across 17 communities"

author: 
  - name          : "Manuel Bohn"
    affiliation   : "1,2,*"
    corresponding : yes    # Define only one corresponding author
    address       : "Universitätsallee 1, 21335 Lüneburg, Germany"
    email         : "manuel.bohn@leuphana.de"
    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
      - "Conceptualization"
      - "Methodology"
      - "Formal Analysis"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Julia Christin Prein"
    affiliation   : "1,2,*"
    role:
      - "Conceptualization"
      - "Methodology"
      - "Software"
      - "Investigation"
      - "Writing - Review & Editing"
  - name          : "Agnes Ayikoru"
    affiliation   : "3"
    role:
      - "Investigation"
  - name          : "Florian M. Bednarski"
    affiliation   : "4"
    role:
      - "Investigation"
      - "Writing - Review & Editing" 
  - name          : "Ardain Dzabatou"
    affiliation   : "5"
    role:
      - "Investigation"
  - name          : "Michael C. Frank"
    affiliation   : "6"
    role:
      - "Investigation"
      - "Writing - Review & Editing"
  - name          : "Annette M. E. Henderson"
    affiliation   : "4"
    role:
      - "Investigation"
      - "Writing - Review & Editing"      
  - name          : "Joan Isabella"
    affiliation   : "3"
    role:
     - "Investigation" 
  - name          : "Josefine Kalbitz"
    affiliation   : "2"
    role:
      - "Investigation"
      - "Writing - Review & Editing"
  - name          : "Patricia Kanngiesser"
    affiliation   : "7"
    role:
      - "Investigation"
      - "Writing - Review & Editing" 
  - name          : "Dilara Keşşafoğlu"
    affiliation   : "8"
    role:
      - "Investigation"
      - "Writing - Review & Editing" 
  - name          : "Bahar Köymen"
    affiliation   : "9"
    role:
      - "Investigation"
      - "Writing - Review & Editing"  
  - name          : "Maira V. Manrique-Hernandez"
    affiliation   : "2"
    role:
      - "Investigation"
  - name          : "Shirley Magazi"
    affiliation   : "10"
    role:
      - "Investigation"
  - name          : "Lizbeth Mújica-Manrique"
    affiliation   : "2"
    role:
      - "Investigation"
      - "Writing - Review & Editing" 
  - name          : "Julia Ohlendorf"
    affiliation   : "2"
    role:
      - "Investigation"
  - name          : "Damilola Olaoba"
    affiliation   : "2"
    role:
     - "Investigation"
  - name          : "Wesley R. Pieters"
    affiliation   : "10"
    role:
      - "Investigation"
      - "Writing - Review & Editing" 
  - name          : "Sarah Pope-Caldwell"
    affiliation   : "2"
    role:
     - "Investigation"    
  - name          : "Katie Slocombe"
    affiliation   : "11"
    role:
     - "Investigation"
     - "Writing - Review & Editing"
  - name          : "Robert Z. Sparks"
    affiliation   : "6"
    role:
      - "Investigation"
  - name          : "Jahnavi Sunderarajan"
    affiliation   : "2"
    role:
      - "Investigation"
  - name          : "Wilson Vieira"
    affiliation   : "2"
    role:
     - "Investigation"
  - name          : "Zhen Zhang"
    affiliation   : "12"
    role:
      - "Investigation"
      - "Writing - Review & Editing"   
  - name          : "Yufei Zong"
    affiliation   : "12"
    role:
      - "Investigation"
  - name          : "Roman Stengelin"
    affiliation   : "2,10,+"
    role:
      - "Conceptualization"
      - "Methodology"
      - "Investigation"
      - "Writing - Review & Editing" 
  - name          : "Daniel B. M. Haun"
    affiliation   : "2,+"
    role:
      - "Conceptualization"
      - "Funding acquisition"
      - "Writing - Review & Editing"

@@

authornote: |
   The authors would like to thank Luke Maurits for statistical advice. Manuel Bohn was supported by a Jacobs Foundation Research Fellowship (2022-1484-00). We are grateful to thank all children and caregivers for participating in the study. We thank the Max Planck Society for the Advancement of Science.

abstract: |
  Theoretical accounts assume that key features of human social cognition are universal. Here we focus on gaze following, the bedrock of social interactions and coordinated activities, to test this claim. In a comprehensive cross-cultural study spanning five continents and 17 distinct cultural communities, we examined the development of gaze following in early childhood. We identified key processing signatures through a computational model that assumes that participants follow an individual’s gaze by estimating a vector emanating from the eye center through the pupil. We found these signatures in all communities, suggesting that children worldwide processed gaze in highly similar ways. Absolute differences between groups were accounted for by a cross-culturally consistent relationship between children’s exposure to touchscreens and their performance in the task. These results provide strong evidence for a universal process underlying a foundational socio-cognitive ability in humans that can be reliably inferred even in the presence of cultural variation in overt behavior.

bibliography      : "library.bib"

@@

classoption       : "man"
csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass     : "apa7"
output            :
  papaja::apa6_pdf:
    latex_engine: xelatex
    keep_tex: true
editor_options    : 
  markdown: 
    wrap: sentence
header-includes:
  - \usepackage{amsmath}
  - \usepackage{amssymb}
  - \usepackage{unicode-math}
  - \usepackage{setspace}
  - \usepackage{libertine}
  - \captionsetup[figure]{font={stretch=1}}
  - |
    \makeatletter
    \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-1em}%
      {\normalfont\normalsize\bfseries\typesectitle}}

    \renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-\z@\relax}%
      {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
    \makeatother
---

@@

# Research Transparency Statement

All authors declare no conflicts of interest. Preregistration: The hypotheses, methods and parts of the analysis plan were preregistered (<!-- https://osf.io/tdsvc/?view_only=f404d66c967542f28ce4af16371f34e6-->https://osf.io/tdsvc) on March 12th, 2022, prior to data collection which began on March 18th, 2022. Additional analysis and deviations from the preregistration are reported in the Supplementary Material. Materials: All study materials are publicly available <!--[masked for peer review] --> (https://ccp-odc.eva.mpg.de/tango-cc/). Data: All primary data are publicly available <!--[masked for peer review; see .zip file included--> (https://github.com/ccp-eva/gafo-cc-analysis/). Analysis scripts: All analysis scripts are publicly available <!--[masked for peer review; see .zip file included]--> (https://github.com/ccp-eva/gafo-cc-analysis/).

# Introduction

Human socio-cognitive skills enable unique forms of communication and cooperation that provide a bedrock for cumulative culture and the formation of complex societies  [@tomasello2020adaptive; @laland2021understanding; @wellman2014making; @henrich2016secret; @tomasello2003makes; @legare2019development; @heyes2018cognitive]. The eyes are the proverbial "window to the mind" and eye gaze is essential for many social reasoning processes  [@shepherd2010following; @doherty2006development; @emery2000eyes]. Others’ eye gaze is used to infer their focus of visual attention, which is a critical aspect of coordinated activities, including communication and cooperation [@tomasello2007reliance; @scaife1975capacity; @langton2000eyes; @rossano2012gaze; @richardson2005looking; @sebanz2006joint].

The ability to follow gaze emerges early in development [@tang2023slow; @gredeback2010development; @byers2021development; @del2019developmental]. The earliest signs of gaze following have been found in infants as young as four months [@astor2021gaze; @d1997demonstration]. Initially, infants rely more on head direction than actual gaze direction [@michel2021effects; @lempers1977development]. Throughout the first two years of life, children refine their abilities: they interpret gaze in mentalistic terms, for example, they follow gaze to locations outside their own visual field by moving around barriers [@moll200412]. Importantly, individual differences in children’s gaze following abilities predict later life outcomes, most notably communicative abilities [@carpenter1998social]. For example, gaze following at 10 months predicts language scores at 18 months of age [@brooks2005development]. Difficulties with gaze following have been linked to developmental disorders, including Autism [@itier2009neural; @thorup2016altered; @thorup2018reduced]. This work highlights the importance of gaze following as a foundational building block of human social interaction and its central place in theorizing.

A central assumption in the theoretical and empirical work discussed above is that, despite substantial variation in developmental contexts, gaze following works and develops in the same way across human societies [@tomasello2019becoming]. This assumption -- despite being central to many developmental theories -- is currently not supported by evidence. On the contrary, cross-cultural studies have revealed substantial diversity in socio-cognitive development [@miller2018contributions; @wellman2014making; @mayer2013synchrony; @dixson2018scaling; @taumoepeau2019cross]. One of the very few cross-cultural studies also found differences in the likelihood to follow gaze between communities [@callaghan2011early].

One potential source for this paradox lies in the reliance on aggregated measures in cross-cultural studies. Absolute differences in mean performance across communities are interpreted as a signal of different underlying cognitive processes. In the present study, we resolve this paradox by instead focusing on processing signatures that can be investigated independently of absolute community-level differences. This allows us to directly evaluate the empirical foundation of claims about universal features of human social cognition. To this end, we conducted a pre-registered, large-scale, cross-cultural study on the development of gaze following abilities to study potentially universal processing signatures.

The processing signatures were derived from a computational model that assumes that participants follow gaze by estimating a vector emanating from the eye center through the pupil [@prein2024variation]. The key innovation of the model is that it explains how individuals may use the same cognitive process but still differ in their measured abilities. The process always involves estimating a vector but also involves a degree of uncertainty because the eye center is not directly observable. Individuals are assumed to differ in their level of uncertainty with which they estimate the vector which causes differences in their observable behavior. Importantly, the assumed process leaves a key signature in the data that is observable independent of the absolute level of performance. In the present study, we therefore focus on this signature instead of absolute levels of performance when evaluating the claim whether there is evidence for a universal cognitive mechanism underlying gaze following.

The `r n_distinct(data$subjid)` participants who took part in the study lived in 17 different communities across 14 countries and five continents (Fig. \@ref(fig:fig1)A, Tab. \@ref(tab:tab1)). These countries represent ~46% of the world’s population. Communities covered a broad spectrum of geographical locations, social and political systems, languages, and subsistence styles (see Supplementary Material). This diversity allowed us to overcome the common pitfall of cross-cultural studies that compare urban communities from the Global North to rural communities from the Global South [@barrett2020towards].

```{r fig1, include = T, fig.align = "center", fig.cap = "(A) Data collection sites. Points show the approximate geographical location of the data collection sites, coloring shows the sample sizes. (B) Screenshots from the task. The upper scene depicts the start, and the lower scene depicts the choice phase in a test trial. Participants had to use the gaze of the agent to locate the balloon and touch the location on the hedge where they thought the balloon was. Agents, audio recordings and backgrounds were adapted to each community. (C) Drawings used as agents across communities.", out.width="100%"}
knitr::include_graphics("../figures/fig1_2.png")
```

```{r tab1, include = T}
tba1 <- data %>%
  separate(community, into = c("community", "country"), sep = " \\(")%>%
  mutate(country = str_remove(country, "\\("),
         country = str_remove(country, "\\)"))%>%
  distinct(subjid, .keep_all = T)%>%
  group_by(community, country, continent)%>%
  summarise(N = n_distinct(subjid), 
            m = sum(sex == "m", na.rm = T), 
            mean_age = mean(ageinyears),
            min_age = min(ageinyears), 
            max_age = max(ageinyears), 
            touchscreen = mean(touchscreen, na.rm = T))%>%
  mutate_if(is.numeric, format, digits = 2, nsmall = 2)%>%
  mutate(age = paste0(mean_age, " (", min_age, " - ", max_age, ")"))%>%
  mutate(N = paste0(N, " (", m,")"))%>%
  select(-c(mean_age, min_age, max_age))%>%
  arrange(continent, country)%>%
  group_by(continent) %>%
  mutate(continent = as.character(continent),
         continent = replace(continent, duplicated(continent), " ")) %>%
  ungroup() %>%
  mutate(country = replace(country, duplicated(country), " "))%>%
  mutate(language = c("Spanish (Rioplatense)", "Spanish (Mexican)", "English (American)", "Hai||om", "Khwedam", "English (Nigerian)2", "English (Nigerian)", "Yaka", "Lingala", "Kiswahili", "Bemba", "German", "English (British)", "Mandarin", "English (Indian) / Marathi", "Turkish", "English (New Zealand)"))%>%
  select(continent, country, community, N, age, language, touchscreen)

apa_table(tba1,
          col.names = linebreak(c("Continent", "Country", "Community", "N (male)", "Age (range)", "Language", "Touchscreen exposure1")),
          caption = "Participant demographics.",
          align = paste0("m{", 1/(ncol(tba1) + 1), "\\linewidth}"),
          longtable = TRUE,
          note = "1 Proportion of participants who have access to touchscreens according to parental questionnaire. 2 Local collaborators and piloting suggested that Nigerian English is suitable for Windhoek as well.",
          escape=TRUE)
```

We used an animated picture book tablet task in which participants had to locate a hidden object based on observing an agent’s gaze. Children watched a balloon disappear behind a hedge. An agent followed the trajectory of the balloon with their eyes (Fig. \@ref(fig:fig1)B). The key dependent variable was the (im)precision with which children located the agent’s focus of attention, that is, the deviation between where the agent looked (where the balloon was) and the child’s response. We adapted visuals and audio instructions specifically for each of the 17 communities. Previous work demonstrated excellent individual-level measurement properties for this task in a German sample [@prein2024tango].

@@

# Methods

## Participants

```{r}
demtab <- data %>%
  separate(community, into = c("community", "country"), sep = " \\(")%>%
  mutate(country = str_remove(country, "\\("),
         country = str_remove(country, "\\)"))%>%
  distinct(subjid, .keep_all = T)%>%
  group_by(community, country, continent)%>%
  summarise(N = n_distinct(subjid), 
            m = sum(sex == "m", na.rm = T), 
            mean_age = mean(ageinyears),
            min_age = min(ageinyears), 
            max_age = max(ageinyears))%>%
  mutate_if(is.numeric, format, digits = 2, nsmall = 2)%>%
  mutate(age = paste(mean_age, " (", min_age, " - ", max_age, ")"))%>%
  select(-c(mean_age, min_age, max_age))%>%
  arrange(continent)
```

A total of `r sum(as.numeric(demtab$N))` children between `r round(min(data$ageinyears), 2)` and `r round(max(data$ageinyears), 2)` years of age provided data for the study. Children lived in `r n_distinct(data$community)` different communities, located in `r n_distinct(demtab$country)` different countries. Table \@ref(tab:tab1) gives the sample size per community together with some basic demographic information. The recruitment strategy for each community is reported in the respective site descriptions. For some children, the exact birthday was unknown. In such cases, we set the birthday to the 30th of June of the year that would make them fall into the reported age category.

Data from children was only included in the study when they contributed at least four valid test trials. We also excluded the data from children with a diagnosed developmental disorder. In sum, in addition to the sample size reported above, 74 additional children participated in the study but did not contribute data. The main reasons for exclusion were: contribution of less than four valid test trials, technical failures, and missing or implausible demographic information (e.g., when the number of children living in the household was reported to be larger than the household itself or when the number of children reported to live in the household equaled the number of children younger than the child being tested). We did not exclude any participants for performance reasons. A detailed description of each data collection site and the way children were recruited can be found in the Supplementary Material.

## Setup and Procedure

The task was implemented as a browser-based interactive picture book using `HTML`, `CSS`, and `JavaScript`. Participants saw animated agents on a touch screen device, listened to pre-recorded audio instructions and responded by touching the screen. In all communities, a research assistant, fluent in the local language(s), guided the child through the task.

Figure \@ref(fig:fig1)B shows a screenshot from the task. The task was introduced verbally by the assistant as the balloon game in which the participant would play with other children to find a balloon. On each trial, participants saw an agent located in a window in the center of the screen. A balloon fell down from its starting position just below the agent. The agent's gaze followed the trajectory of the balloon. That is, the pupils and the iris were programmed to align with the center of the balloon. Once the balloon had landed on the ground, the agent was instructed to locate it, that is, to touch the location on the screen where they thought the balloon was. On each trial, we recorded the exact x-coordinate of the participant's touch.

There were two types of training trials. In training 1 trials, the balloon fell down and landed in plain sight. Participants simply had to touch the visible balloon. In training 2 trials, the trajectory of the balloon was visible but it landed behind a small barrier (a hedge -- see Figure \@ref(fig:fig1)B). Thus, participants needed to touch the hedge where they saw the balloon land. Next came test trials. Here, the barrier moved up and covered the balloon's trajectory. That is, participants only saw the agent's eyes move, but not the balloon. They had to infer the location of the balloon based on the agent's gaze direction. During training 1, training 2 and the first test trial, children heard voice-overs commenting what happened on the screen. Critically, the agent was described as wanting to help the child and always looking at the balloon.

Children completed one training 1, two training 2 trials and 16 test trials. We excluded the first test trial from the analysis because of the voice-over. Thus, 15 test trials were used in the analysis below. Each child saw eight different agents (four male, four female). The agent changed from trial to trial, with alternating genders. A coin toss before the first trial decided whether the first agent was male or female. The order in which agents were shown was randomized with the constraint that all agents had to be shown once until an agent was shown again. The color of the balloon also changed from trial to trial in a random order, also with the constraint that all colors appeared once before any one was repeated.

The location (x-coordinate) where the balloon landed was determined in the following way: The screen was divided in ten equally sized bins. On each trial, one of the bins was randomly selected and the exact x-coordinate was randomly chosen within that bin. Constraints were that the balloon landed in each bin equally often and the same bin appeared no more than twice in a row.

All children were tested with a touchscreen device with a size between 11 and 13 inch equipped with a webcam. The data was either stored locally or sent to a server. In addition to the behavioral data, we stored the webcam recording of the session for verification purposes. Community-specific adaptations were made by changing the visuals and the audio instructions (see Supplementary Material for details).

In addition to the gaze following task, caregivers responded to a short questionnaire about children’s access to screens and touchscreens (binary answer) as well as the number of people, children and children younger than the focal child living in the household (numeric; see Supplementary Material for details). The numeric variables were scaled within community prior to inclusion into the regression models.

# Analysis

## Regression models

We used Bayesian Regression models fit in `R` [@r] using the package `brms` [@burkner2017brms] for all analyses except the cognitive models (see below). We used default priors built into `brms`. The dependent variable in all regression models was imprecision, that is, the absolute distance between the true location of the balloon (x-coordinate of its center) and the location where the participant touched the screen. We used a Log-normal distribution to model the data because the natural lower bound for imprecision is zero and the data was right skewed with a long tail. Numeric predictors that entered the models were scaled to have a mean of zero and a standard deviation of 1.

To analyse cross-cultural variation in performance, we used a cross-validation procedure [see e.g., @stengelin2023children]. In the Supplementary Material we give a detailed justification of this approach. For each community, we randomly sampled a data set that was 5/6 the size of the full data set (training data). Then, we fit the model to this training data and used the estimated model parameters to predict the remaining 1/6 of the data (testing data). We then compared the model predictions from the different models by computing the mean difference between the true and predicted imprecision, over all trials in the testing data set. We repeated the cross-validation procedure 100 times and computed the percentage of cases in which one model outperformed the other. We compared three models: a null model assuming no systematic community-level variation, a model assuming variation between communities and a model assuming variation between communities and in developmental trajectories (see Supplementary Material for model equations).

To evaluate the processing signatures predicted by the cognitive model that trials in which the balloon lands further away from the center lead to larger imprecision (see next section for details), we fit a model predicting imprecision by age and target centrality (distance of the landing position from the center in pixel/SVG units) with random intercepts for participant and community and random slopes for target centrality within participant and community (`brms` notation: `age + target_centrality + (target_centrality | participant) + (age + target_centrality | community)`).

## Cognitive model

Recent computational work modeled gaze following as social vector estimation [@prein2024variation]. When following gaze, onlookers observe the location of the pupil within the eye and estimate a vector emanating from the center of the eye through the pupil. The focus of attention is the location where the estimated vectors from both eyes hit a surface (Fig. \@ref(fig:fig12)). It is assumed that this estimation process has some uncertainty because the center of the eye is not directly observable and that individuals vary in their level of uncertainty. As a consequence, even though individuals use the same general process, they might differ in their absolute levels of precision. Crucially, this process model predicts a clear performance signature in the gaze following task: Trials in which the agent looks further away from the center should result in lower levels of precision compared to trials in which the agent looks closer to the center. This prediction is best understood by considering a similar phenomenon: pointing a torch light to a flat surface. The width of the light beam represents each individual’s level of uncertainty in vector estimation. When the torch is directed straight down, the light beam is concentrated in a relatively small area. When the torch is rotated to the side, the light from one half of the cone must travel further than the light from the other half to reach the surface. As a consequence, the light is spread over a wider area (see Fig. \@ref(fig:fig12)).

The model inversely models the process generating touches on the screen based on observed eye movements and is defined as:

\begin{equation}
    P(\theta | x_c, \alpha_l, \alpha_r) \propto P(x_c | \alpha_l, \alpha_r, \theta)P(\theta)
\end{equation}

Here, $\theta$ represents an individual's cognitive ability to locate the focus of the agent's attention, $x_c$ represents the touched coordinate, and $\alpha_l$ and $\alpha_r$ correspond to the left and right pupil angles (each defined as the angle between a line connecting the center of the eye to the pupil and a line extended vertically downward from the center of the eye).

The basic assumption in this model is that participants touch on the screen location where they think the agent is looking. The true eye angles ($\alpha_l$ and $\alpha_r$) are not directly observable and are estimated with noise, yielding $\hat{\alpha_l}$ and $\hat{\alpha_r}$.

Each touch $x_c$ implies a "matched pair" of estimated pupil angles $\hat{\alpha_l}$ and $\hat{\alpha_r}$, with the constraint that the lines extended along those two angles meet at the precise location of where the target is believed to be. As a consequence, we can rewrite the likelihood function of the model as:

\begin{equation}
P(x_c | \alpha_l, \alpha_r, \theta) \propto P(\hat{\alpha_l}, \hat{\alpha_r} | \alpha_l, \alpha_r, \theta) P(x_c)
\end{equation}

$P(x_c)$ is a prior over potential target locations. Because the target was last visible in the screen and because the agent was located in the center, we assumed that participants have an a priori expectation that the target will land close to the middle. We estimated the strength of this center bias (*i.e.,* the standard deviation of a Normal distribution around the screen center) based on the data: $P(x_c) \sim \mathcal{N}(960, \sigma^p)$.
